---
title: "ANOVA analysis of frog dataset"
author: "Thomas M Smith"
date: "2024-02-27"
output:
  pdf_document: default
  html_document: default
  latex_engine: pdflatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.width = 8,
                      fig.asp = 0.8,
                      out.width = "80%")

```

```{r library, include=FALSE}
# Packages ----
library(tidyverse)
library(janitor)
library(gt)
```

```{r read-data, include=FALSE, echo=FALSE}
#___________________----
# Read data----
frog_data <- read.csv("data/frogs_messy_data.csv")
#_______________----
# Clean data ----
# clean up column names
frog_data_clean <- janitor::clean_names(frog_data) %>%
  rename("13" = temperature13,
         "18" = temperature18,
         "25" = temperature25,
         frogspawn_id = `frogspawn_sample_id`) %>% 
  pivot_longer(`13`:`25`, names_to="temperature", values_to="days") %>% 
  drop_na(days)
# check data is in a tidy format
head(frog_data_clean)
# check variable names
colnames(frog_data_clean)
# check for duplication
frog_data_clean %>% 
  duplicated() %>% 
  sum()
# check for typos - by looking at impossible values
frog_data_clean %>% 
  summarise(min=min(temperature, na.rm=TRUE), 
            max=max(temperature, na.rm=TRUE))
frog_data_clean %>% 
  summarise(min=min(days, na.rm=TRUE), 
            max=max(days, na.rm=TRUE))
# check for typos by looking at distinct characters/values
frog_data_clean %>% 
  distinct(temperature)
frog_data_clean %>% 
  distinct(days)
# missing values
frog_data_clean %>% 
  is.na() %>% 
  sum()
# quick summary
summary(frog_data_clean)
```
# Results from ANOVA
### Raw and tabled data
```{r analysis, include=TRUE, echo=FALSE, warning=FALSE}
#_____________________----
# Analysis----
# ANALYSIS ----
lsmodel_frogs <- lm(days ~ temperature, data = frog_data_clean)
summary(lsmodel_frogs)
anova(lsmodel_frogs)%>% gt::gt()
broom::tidy(lsmodel_frogs, conf.int = T)%>% gt::gt()
```
# Putting the ANOVA test into words
### Increasing enivornmental temperatures had a clear negative effect on the time taken for frogspawn to hatch (one-way ANOVA: F2,57 = 385.9, P < 0.001). At 13C the mean time to hatching was 26.3 days, this decreased by an average of 5.3 days at 18C and finally by 10.1 days at 25C.
## Visualised dataset (Figure 1) previously analysed using ANOVA
```{r plotting, include=TRUE, echo=FALSE, warning=FALSE}
#___________________________----
# Plotting ----

boxpplot_frog <- ggplot(data=frog_data_clean, aes(x=temperature, y=days))+
  geom_boxplot(aes(fill=temperature),
               alpha = 0.5,
               width = 0.425)+
  geom_jitter(aes(colour=temperature),
              width=0.2)+
  labs(x="Temperature (°C)", y="No. of days for spawn to hatch",
       title = "Figure 1", subtitle = "Hatching times in days respective to three differing Temperatures (°C)")+
  scale_fill_manual(values = c("cyan2", "coral1","darkred"))+
  scale_colour_manual(values = c("cadetblue4", "coral3","brown4"))+
  theme_minimal()+
  theme(legend.position = "none")

boxpplot_frog

```
### Figure 1: The time in days it takes for frogspawn to hatch in 13, 18 and 25C temperatures. Thick horizontal line is the median with individual points displayed as a jitter.   